---
title: "Vignette template"
author: "Maxime Jaunatre"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{template}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r opts, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  # include = FALSE, # don't evaluate code chunk
                      eval = FALSE, # don't run code
                      echo = FALSE, # hide code
                      message = FALSE, # hide message provoked by code
                      warning = FALSE, # idem but for warningss
                      fig.cap = "..." # don't know what it does
)
```

```{r setup}
library(BeePODYNA)
```

Ceci est un test de vignette pour voir un peu comment faire. package `BeePODYNA`


# Heading 1
## Heading 2
### Heading 3

--------
********

* Bulleted list
* Item 2
    * Nested bullets need a 4-space indent.
    * Item 2b
    
      * It's possible to put multiple paragraphs of text in a list item. 

    But to do that, the second and subsequent paragraphs must be
    indented by four or more spaces. It looks better if the first
    bullet is also indented.
    
1. Item 1
1. Item 2
1. Items are numbered automatically, even though they all start with 1.

Definition
  : a statement of the exact meaning of a word, especially in a dictionary.

List 
  : a number of connected items or names written or printed consecutively, 
    typically one below the other. 
  : barriers enclosing an area for a jousting tournament.
  
_italic_ or *italic*
__bold__ or **bold**    
[link text](destination)
<http://this-is-a-raw-url.com>


| Right | Left | Default | Center |
|------:|:-----|---------|:------:|
|   12  |  12  |    12   |    12  |
|  123  |  123 |   123   |   123  |
|    1  |    1 |     1   |     1  |

```{r, results = "asis"}
pander::pandoc.table(iris[1:3, 1:4])
```



__A vignette is a long-form guide to your package. Function documentation is great if you know the name of the function you need, but it’s useless otherwise. A vignette is like a book chapter or an academic paper: it can describe the problem that your package is designed to solve, and then show the reader how to solve it. A vignette should divide functions into useful categories, and demonstrate how to coordinate multiple functions to solve problems. Vignettes are also useful if you want to explain the details of your package. For example, if you have implemented a complex statistical algorithm, you might want to describe all the details in a vignette so that users of your package can understand what’s going on under the hood, and be confident that you’ve implemented the algorithm correctly.__

__You’ll need to watch the file size. If you include a lot of graphics, it’s easy to create a very large file. There are no hard and fast rules, but if you have a very large vignette be prepared to either justify the file size, or to make it smaller.__

