---
title: "Dive Planification"
author: "Maxime Jaunatre"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Dive Planification}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---



```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.height = 6.1,
  fig.width = 8.5,
  dpi = 96
)
#library(details)
library(viridisLite)
```

The package `{DiveR}` provide functions to simulate a dive. It does so by asking a depth, a time and few parameters. This vignette will demonstrate the workflow to simulate a simple dive and different models of two-dives systems.

**Note here that it only show square dives, meaning that decompression is managed following mn90 (french) dive tables.**

## Simple Dive

To simulate a simple dive using tables, you just need to provide a depth (meter) and a time (minute) at this depth. Despite being an altitude below 0, the depth must be set in positive values in the package.

```{r basic_dive}
library(DiveR)
dive20_40 <- dive(depth = 20, time = 40)
```

This object of class dive contain a time/depth curve, information about the desaturation process and more information about the parameters used and other results that will be used in plots or other simulations.

The default parameters worth noting are :
* `secu = TRUE` adding a default security stop at 3m for 3min
*`desat_model` The desaturation model used. This vignette will only show `table` model (mn90 french dive tables).

### Limits of table model

This precise model provide desaturation stops at 3, 6 and 9 meters depths for a given time and depth. It's important to check these two values because this table is limited to a maximum depth of 65m and time limits depending on the depth. 

```{r secu_curve,  dev='png', out.width="100%", echo = FALSE}
nodeco <- time <- depth <- 1:65
for(i in depth){
  time[i] <- max_depth_time(depth[i])
  nodeco[i] <- max_depth_time(depth[i], force = TRUE, no_deco = TRUE)
}
cols = viridis(3)
# plot(time, -depth, type = "l", xlim = c(0, max(time)), 
#      xlab = "Time (min)", ylab = "Depth (meter)", 
#      main = "Maximum time in table with desat stop (Deco) and without")
# lines(nodeco, -depth, type = "l")
# polygon(c(rep(0, length(nodeco)), rev(nodeco)), c(-depth, -rev(depth)), col = cols[2])
# polygon(c(time, rev(nodeco)), c(-depth, -rev(depth)), col = cols[1])

df <- rbind(rev(nodeco),rev(time-nodeco))
colnames(df) <- -rev(depth)
barplot(df, horiz = TRUE, col = rev(cols[-3]), yaxt = "n", 
     xlab = "Time (min)", ylab = "Depth (meter)", 
     main = "Maximum time in table with desat stop (Deco) and without")
axis(2, at = seq(0, par("usr")[4]- 3, length.out = 14), labels = seq(-65, -0, by = 5))
legend("bottomright", legend = c("Deco", "No-deco time"), fill = (cols[-3]))
```


```{r lim funct, error=TRUE}
### check if in table
tablecheck(20, 50)
tablecheck(20, 80)
tablecheck(20, 80, force = TRUE)
### checking time
max_depth_time(depth = 20)
max_depth_time(depth = 20, no_deco = TRUE)
```

<!--
<details>
 <summary>Summary Goes Here</summary>
 ...this is hidden, collapsable content...
</details>


## Dive Methods

## Multiple Dives

### Independant Dives

### Successive Dives

### Consecutive Dives


```{r basic_tank}
library(DiveR)
tank1 <- tank(vol = 12, press = 200)
tank1
```
-->
