---
title: "Gas consumption"
author: "Maxime Jaunatre"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Gas consumption}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The package `{DiveR}` provide functions to simulate gas consumption. It does so by using tank object and putting them in link with a dive curve and few parameters. This vignette will demonstrate the workflow to use tanks, for a dive with a single tank to using multiples ones on different setups.

## Tank creation

A tank can basically be set by 2 important parameters : a volume `vol` and a pressure `press`. Other input are used to set rules to define when and where the tank can be used to breath.

Volume is set in litre and pressure in bar. 

```{r basic_tank}
library(DiveR)
tank1 <- tank(vol = 12, press = 200)
tank1
```

We can see that many values are set by default, the principal one being the gas type. At this time only air is used and thus tanks are limited between the surface (0) and 66m of depth.

Default rules are also set, which will message the user when the tank is half empty, 3/4 empty and when there is no gas left in it. You can set them like below, and name the rules too. 

```{r limit_tank}
# set in percentage
tank2 <- tank(vol = 12, press = 200, 
              rules = list(rules = c('mid' = 50, 'quarter' = 25), sys = '%'))
tank2$carac[c('rule1', 'rule2')] # values
tank2$typo[c('rule1', 'rule2')] # names
# set in bar
tank3 <- tank(vol = 12, press = 200, 
              rules = list(rules = c('return' = 130, 'end' = 80), sys = 'bar'))
tank3$carac[c('rule1', 'rule2')] # values
tank3$typo[c('rule1', 'rule2')] # names
```

### Tank types

By default, the tank type is set to *back*, a category to describe back-mounted tanks commonly used by divers. However it is possible to have tanks dedicated to specific usage in a relay or at the end of a dive. 

#### Relay tank

A tank used at the beginning and the end of a dive. Mostly used in cave diving. This type is attributed to tanks where there is an interdiction of usage between 2 times.

The easier way to input this interdiction is to set the two rules at same value. During the dive, the tank will be used until pressure match 120b. Then the diver will use other tanks up to their second rules (ex : back tank reserve at 50b). When done, the diver will get back to the relay tank to finish the dive with it. 

```{r relay_tank}
relay <- tank(vol = 12, press = 180, typ = 'relay',
              rules = list(rules = c('gap' = 120, 'end_gap' = 120), sys = 'bar'))
relay$carac[c('rule1', 'rule2')]
```
<!-- Other gas not yet included.

#### Desat tank

A tank used at the end of a dive to desaturate azote during ascent part of the dive. Often limited to a lower depth because it use other gas than air. This type of tank are tanks that divers can't use until he has used other tanks or passed the dive end point.

You can add an interdiction of using a tank by setting the first rule with the same value as the tank pressure.

```{r desat_tank}
desat <- tank(vol = 12, press = 180, typ = 'relay',
              rules = list(rules = c('gap' = 100, 'end_gap' = 25), sys = '%'))
desat$carac[c('rule1', 'rule2')]
```
-->

As these tanks are used in combination with a back-mounted tank, their usage will be described in the multiple tank usage part. The algorithm of usage will be described, don't fear not understanding it right now.


## Tank attributes

The parameters of a tank can be retrieved with specified function. We can find the pressure of any tank.

```{r attribute_tank}
pressure(relay)
volume(relay)
```

<!-- 
```{r tmp_attribute_tank, include = FALSE}
rules <- function(x){

}

summary_conso <- function(object, ...){
  # parameters volume and pressure
  cat(paste("Tank :",volume(object), "litre at",pressure(object),"bar\n"))
  # dtr
  dtr <- dtr(object)
  cat('Total dive time is',dtime(object)+dtr,'with a dive ascent of',dtr,'minutes\n')
  # palier & maj
}

```


--> 

## Tank consumption

Once the tanks are defined, you can combine them with a dive. The `conso` function will then compute how many gas the diver will consume by following the dive depth curve and report pressure at different times of the dive. The function will also catch times where the rules of a tank are met. Note here that a new rules about air failure is added to each tank to report when a tank is empty. 

If no tanks are available (pressure > 0) at some time, the function will return warnings to highlight these times.

For the following examples, we will use a single dive of 40 minutes at 20 meter. There is one desaturation stop for 3min at 3m.

```{r ex_dive}
simp_dive <- dive(depth = 20, time = 40, secu = TRUE)
summary(simp_dive)
```

For mutliple dives you need to choose one of the two dives. A special case for consecutive dives is at work.

### Single Tank

First we try to use a tank with a volume of 12L at 200 bar. The diver will breathe 20L/min, which is the default value of the function. 

A rapid observation of the consumption if possible by ploting the consumption object. We can easily see when the pressure met the rules and when the diver died (40min). This was also printed in warnings with specified times with missing gas.

```{r death_conso}
Tank_12L <- tank(vol = 12, press = 200)
death <- conso(dive = simp_dive, tank = Tank_12L, 
               cons = 20, failure_label = 'Air failure')
plot(death)
```
Once we observed this, we can solve the issue by using a different tank for example. Here we will increase the volume to 15L to get more air during the dive. Note that we could also increase the pressure of the precedent tank. Another solution could be to breath less and using a diver using  `cons = 15`.

This solution work and the diver get to the end of the dive alive ! Hourra ! 

```{r viable_conso}
Tank_15L <- tank(vol = 15, press = 200)
viable <- conso(dive = simp_dive, tank = Tank_15L, 
               cons = 20, failure_label = 'Air failure')
plot(viable)
```

## Consumption attributes

Now that we have simulated both consumption on the same dive, we may want to find more information about the simulations, which will help us dive safely. 

First we can check the pressure at the end of the dive. This can be helpful to see if the diver can still be alive. Here we see that effectively in the death simulation, the tank was empty at the end of the dive.

```{r pressure_conso}
pressure(death)
```

<!-- 

However this does not tell us when this happened. To find out when, we can check when the rules have been met with the `rules` function. This will help us check if the dive instructor interdiction to use the reserve of the tank (here 50bar) in both simulations.

```{r rules_conso, include = FALSE}

rules <- function(conso, tankn, n){
  # check if any rules are crossed for every tank.
  # tank n  being the number of the tank used, by default 0 will show all the tanks
  # n is the number of the rule 1 rule1, 2 rule2, 3 AF, 0 all (default), 
  # print('hello')
}

rules(death)
rules(viable)

```

Finally we can find all information in a summary. 

```{r sum
_conso, include = FALSE}

summary_conso <- function(conso){
  # print('hello')
}

summary(death)
summary(viable)

```
--> 

